A numerical study of the solution of X-Mode equations around the hybrid resonance

### 2015-09-15 report_clean_final/TEX/timedomain.tex
Correcting typoes as asked by the referee
fixed multiple typos extra

\section{Time-Dependent Problem Discretization}
\arev{
We develop hereafter the discretization of the time-dependent one-dimensional model (\ref{eq:main_model}), for the particular case $\theta=0$. 
We first show how this system can be discretized thanks to a Yee scheme suggested in \cite{stable_yee_plasma_current}, then a scheme based on semi-lagrangian discretization is employed.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A staggered scheme}
\label{sec:staggered_discretization}
%In this section, we develop a discretization of the time-dependent problem~\eqref{eq:main_model} in one dimension. %In 1D, the derivatives with respect to $y$ vanish, and we are left with the following problem, 
% \begin{align}
% \epsilon_0\partial_t E_{x}=-eN_e u_x, \nonumber\\
% \epsilon_0\partial_t E_{y}+\partial_x H_z=-eN_e u_y,\nonumber\\
% \mu_0\partial_t H_z+\partial_x E_y-\partial_y E_x=0, \label{eq:tdpb}\\
% m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\nonumber\\
% m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y.\nonumber
% \end{align}
We compute a numerical approximation of~\eqref{eq:main_model} \mrev{with $\theta=0$} on a cartesian grid using the Yee scheme, 
as suggested in \cite{stable_yee_plasma_current}. Let $n$ and $\Delta t$ denote the time step index and the time step, as usual the discrete time is $t_n = n \Delta t$. 
It is standard (see \cite{Xu_2006,stable_yee_plasma_current }) to discretize the electrical field on the integer time steps, 
whereas the field $H_z$ on half time steps, $t_{n+1/2}$. The time discretization of the system reads:
 \bealn
 \left\{\begin{array}{l}
 \epsilon_0\frac{ E_{x}^{n+1}-E_x^{n}}{\Delta t}=-eN_e \frac{u_x^{n+1}+u_x^{n}}{2}, \nonumber\\
 \epsilon_0\frac{ E_{y}^{n+1}-E_y^{n}}{\Delta t}+\partial_x H_z^{n+1/2}=-eN_e \frac{u_y^{n+1}+u_{y}^{n}}{2},\nonumber\\
 \mu_0\frac{ H_z^{n+1/2}-H_z^{n-1/2}}{\Delta t}+\partial_x E_y^{n}-\partial_y E_x^{n}=0, \label{eq:tdiscretizedpb}\\
 m_e\frac{ u_{x}^{n+1}-u_x^{n}}{\Delta t}=e\frac{E_x^{n+1}+E_x^{n}}{2}+e\frac{u_y^{n+1}+u_{y}^{n}}{2}B_0-\nu m_e \frac{u_x^{n+1}+u_x^{n}}{2},\nonumber\\
 m_e\frac{u_{y}^{n+1}-u_y^{n}}{\Delta t}=e\frac{E_y^{n+1}+E_y^{n}}{2}-e\frac{u_x^{n+1}+u_x^{n}}{2}B_0-\nu m_e \frac{u_y^{n+1}+u_y^{n}}{2}.\nonumber
\end{array}\right.
 \eealn
In Yee scheme, the different components of the vector fields 
are defined on staggered points. 
We summarize this discretization on Fig.~\ref{schemepos} where the positions 
of \urev{$E_x$}, $u_x$ (black dots), $E_y$, $u_y$ (blue dots) and $H_z$ (red dots) 
on the discretized grid in space and time are displayed. The fully discretized system then reads

\ben 
\eps_0 \frac{E_x\inp- E_x\midin}{\Delta t} = -e N_e \frac{u_x\inp + u_x \midin}{2},
\label{eq:ns1}
\een 
\ben
\eps_0 \frac{E_y\ihnp-E_y\ihn}{\Delta t}+ \frac{H_z\ipnh-H_z\inh}{\Delta x} = -eN_x \frac{u_y\ihnp + u_y\ihn}{2} ,
\label{eq:ns2}
\een
\ben
 \frac{H_z\inh-H_z\inmh}{\Delta t}+ \frac{E_y\ihn-E_y\imhn}{\Delta x}=0,
 \label{eq:ns3}
\een
\ben 
m_e \frac{u_x\inp-u_x\midin }{\Delta x} = e N_e \frac{E_x\inp+E_x\midin}{2}-\nu m_e\frac{u_x\inp+u_x \midin}{2}+ eB_0\frac{u_y\ihn+u_y\ihnp}{2},
\label{eq:ns4}
\een
\ben
m_e \frac{u_y\ihnp-u_y\ihn}{\Delta t} = eN_e \frac{E_y\ihnp + E_y\ihn}{2}- \nu m_e \frac{u_y\ihnp +u_y\ihn}{2}-eB_0 \frac{u_x\midin+u_x\inp}{2},
\label{eq:ns401}
\een
%\input{grid}
\begin{figure}[h!]
	\begin{center}
 \scalebox{1.2}{
		\begin{tikzpicture}

@@

- The first step is to solve 
$$
\left\{
\begin{array}{lll}
\epsilon_0\partial_t E_{x}=0,\\
\epsilon_0\partial_t E_{y}+\partial_x H_z=0,\\
\mu_0\partial_t H_z+\partial_x E_y=0,\\
m_e\partial_t u_x=0,\\
m_e\partial_t u_y=0.
\end{array}
\right.
$$
Since  $\epsilon_0 \mu_0c^2=1$, this can be  performed solving
two transport equations 
$$
\left\{
\begin{array}{lll}
\partial_t w +c \partial_x w=0, & w=E_{y}+ c H_z,  \\
\partial_t z -c \partial_x z=0& z=E_{y}-  c H_z.
\end{array}
\right.
$$
In our case we used 
a semi-lagrangian scheme for a time step $\Delta t$.
In the tests, we used Strang's stencils described in \cite{cedar} with CFL $\nu=0.5$, at order 7. 
It  yields to the chain 
$(E_y,H_z)_i^n\mapsto ( w,z)_i^n\mapsto ( w,z)_i^{n*}\mapsto (E_y,H_z)_i^{n*} $.
Here $n*$ denotes the intermediate time step of the algorithm.

>>>>

\section{Time-Dependent Problem Discretization}
We develop hereafter the discretization of the time-dependent one-dimensional model (\ref{eq:main_model}), for the particular case $\theta=0$. 
We first show how this system can be discretized thanks to a Yee scheme suggested in \cite{stable_yee_plasma_current}, then a scheme based on semi-lagrangian discretization is employed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A staggered scheme}
\label{sec:staggered_discretization}
%In this section, we develop a discretization of the time-dependent problem~\eqref{eq:main_model} in one dimension. %In 1D, the derivatives with respect to $y$ vanish, and we are left with the following problem, 
% \begin{align}
% \epsilon_0\partial_t E_{x}=-eN_e u_x, \nonumber\\
% \epsilon_0\partial_t E_{y}+\partial_x H_z=-eN_e u_y,\nonumber\\
% \mu_0\partial_t H_z+\partial_x E_y-\partial_y E_x=0, \label{eq:tdpb}\\
% m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\nonumber\\
% m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y.\nonumber
% \end{align}
We compute a numerical approximation of~\eqref{eq:main_model} with $\theta=0$ on a cartesian grid using the Yee scheme, 
as suggested in \cite{stable_yee_plasma_current}. Let $n$ and $\Delta t$ denote the time step index and the time step, as usual the discrete time is $t_n = n \Delta t$. 
It is standard (see \cite{Xu_2006,stable_yee_plasma_current }) to discretize the electrical field on the integer time steps, 
whereas the field $H_z$ on half time steps, $t_{n+1/2}$. The time discretization of the system reads:
 \bealn
 \left\{\begin{array}{l}
 \epsilon_0\frac{ E_{x}^{n+1}-E_x^{n}}{\Delta t}=-eN_e \frac{u_x^{n+1}+u_x^{n}}{2}, \nonumber\\
 \epsilon_0\frac{ E_{y}^{n+1}-E_y^{n}}{\Delta t}+\partial_x H_z^{n+1/2}=-eN_e \frac{u_y^{n+1}+u_{y}^{n}}{2},\nonumber\\
 \lastrev{\mu_0\frac{ H_z^{n+1/2}-H_z^{n-1/2}}{\Delta t}+\partial_x E_y^{n}=0,} \label{eq:tdiscretizedpb}\\
 m_e\frac{ u_{x}^{n+1}-u_x^{n}}{\Delta t}=e\frac{E_x^{n+1}+E_x^{n}}{2}+e\frac{u_y^{n+1}+u_{y}^{n}}{2}B_0-\nu m_e \frac{u_x^{n+1}+u_x^{n}}{2},\nonumber\\
 m_e\frac{u_{y}^{n+1}-u_y^{n}}{\Delta t}=e\frac{E_y^{n+1}+E_y^{n}}{2}-e\frac{u_x^{n+1}+u_x^{n}}{2}B_0-\nu m_e \frac{u_y^{n+1}+u_y^{n}}{2}.\nonumber
\end{array}\right.
 \eealn
In \lastrev{the} Yee scheme, the different components of the vector fields 
are defined on staggered points. 
We summarize this discretization on Fig.~\ref{schemepos} where the positions 
of $E_x$, $u_x$ (black dots), $E_y$, $u_y$ (blue dots) and $H_z$ (red dots) 
on the discretized grid in space and time are displayed. The fully discretized system then reads

\ben 
\eps_0 \frac{E_x\inp- E_x\midin}{\Delta t} = -e N_e \frac{u_x\inp + u_x \midin}{2},
\label{eq:ns1}
\een 
\ben
\eps_0 \frac{E_y\ihnp-E_y\ihn}{\Delta t}+ \frac{H_z\ipnh-H_z\inh}{\Delta x} = -e\lastrev{N_e} \frac{u_y\ihnp + u_y\ihn}{2} ,
\label{eq:ns2}
\een
\ben
\lastrev{\mu_0} \frac{H_z\inh-H_z\inmh}{\Delta t}+ \frac{E_y\ihn-E_y\imhn}{\Delta x}=0,
 \label{eq:ns3}
\een
\ben 
m_e \frac{u_x\inp-u_x\midin }{\lastrev{\Delta t}} = \lastrev{e} \frac{E_x\inp+E_x\midin}{2}-\nu m_e\frac{u_x\inp+u_x \midin}{2}+ eB_0\frac{u_y\ihn+u_y\ihnp}{2},
\label{eq:ns4}
\een
\ben
m_e \frac{u_y\ihnp-u_y\ihn}{\Delta t} = \lastrev{e}\frac{E_y\ihnp + E_y\ihn}{2}- \nu m_e \frac{u_y\ihnp +u_y\ihn}{2}-eB_0 \frac{u_x\midin+u_x\inp}{2},
\label{eq:ns401}
\een
%\input{grid}
\begin{figure}[h!]
	\begin{center}
 \scalebox{1.2}{
		\begin{tikzpicture}

@@

- The first step is to solve 
$$
\left\{
\begin{array}{lll}
\epsilon_0\partial_t E_{x}=0,\\
\epsilon_0\partial_t E_{y}+\partial_x H_z=0,\\
\mu_0\partial_t H_z+\partial_x E_y=0,\\
m_e\partial_t u_x=0,\\
m_e\partial_t u_y=0.
\end{array}
\right.
$$
Since  $\epsilon_0 \mu_0c^2=1$, this can be  performed solving
two transport equations 
$$
\left\{
\begin{array}{lll}
\partial_t w +c \partial_x w=0, & w=E_{y}+ c H_z,  \\
\partial_t z -c \partial_x z=0& z=E_{y}-  c H_z.
\end{array}
\right.
$$
In our case we used 
a semi-lagrangian scheme for a time step $\Delta t$.
In the tests, we used Strang's stencils described in \cite{cedar} with CFL $\nu=0.5$, at order 7. 
It  yields  the chain 
$(E_y,H_z)_i^n\mapsto ( w,z)_i^n\mapsto ( w,z)_i^{n*}\mapsto (E_y,H_z)_i^{n*} $.
Here $n*$ denotes the intermediate time step of the algorithm.
